---
title: "4a. supp_RaceEthGenImg_woMCI_02 12 25"
output: html_document
date: "2025-02-13"
---
# Purpose of code
To run main models excluding those with MCI as a sensitivity analysis. 

# Loading packages
```{r packagenews, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(haven) # inputting data 
library(tidyverse) # data mgmt
library(psych) # easy summary statistics
library(DT) # data tables
library(tableone) # easy table 1 
library(kableExtra) # format kable objects 
library(naniar) # for missingness 
library(nnet) # multinomial regression
library(openxlsx) # excel for outputting results 
library(patchwork) # putting plots together
library(RColorBrewer) # color pallettes
options(max.print=100000)

# turn off scientific notation
options(scipen = 999)
```
# Loading data
```{r data, include = FALSE}
load("./aim2_final.RData") # n=3840, v=338
view(names(aim2_final))

# those MCI- and dementia-free at baseline with any imaging data at v1
withimg_nl <- aim2_final %>%
  filter(!is.na(cdx_cog) & cdx_cog == 0) %>% # excluding those w/ MCI and dementia, only cog normal people
  filter( (hasabv1 == 1) | (hasctv1 == 1) | (haswmhv1 == 1) | (hasmttauv1 == 1) | (hashipp1 == 1)) 
 # n=2670, v=338 

# those with amyloid pet data at v1
withabv1_nl <- withimg_nl %>%
  filter(hasabv1 == 1) # n=1584, v=338

# those with tau data at v1
withmttau1_nl <- withimg_nl %>%
  filter(hasmttauv1 == 1) # n=650, v=338

# those with cortical thickness at v1
withctv1_nl <- withimg_nl %>%
  filter(hasctv1 == 1) # n=3032, v=338

# those with hipp vol at v1
withhipp1_nl <- withimg_nl %>%
  filter(hashipp1 == 1) # n=2426, v=338

# those with wmhv data at v1
withwmhv1_nl <- withimg_nl %>%
  filter(haswmhv1 == 1) %>%
  filter(excludeforwmh == 0) # n=2582, v=338

# those with lacunar infarcts at v1
withlac1_nl <- withimg_nl %>%
  filter(!is.na(lacune_1)) # n=2634, v=338

# those with microbleeds at v1
withmb1_nl <- withimg_nl %>%
  filter(!is.na(microhem_1)) # n=2634, v=338

# men
menwithimg_nl <- withimg_nl %>%
  filter(gender == 0) #n=902, v=338

# women
womenwithimg_nl <- withimg_nl %>%
  filter(gender == 1) # n=1768, v=338

```
## Creating excel sheet for results
```{r amyloid excel, include = FALSE}
# if starting over and re-running all models, run this
cognl_wb <- createWorkbook()

# if coming back to analysis, run this
cognl_wb <- loadWorkbook("./Results - RaceEthGenImg/cognl.xlsx")
```

# Models for amyloid deposition 
## Creating gender stratified datasets
```{r amyloid strat data, include = FALSE}
menwithabv_nl <- withabv1_nl %>% filter(gender == 0) # n=356
womenwithabv_nl <- withabv1_nl %>% filter(gender == 1) # n=790
```
## Model 2 with interaction term
```{r amyloid model 2 with intx, include = FALSE}
model2_amyloid_nl <- lm(scale(absuvr_1) ~
                          as.factor(ethnicity) +
                          as.factor(gender) +
                          as.factor(ethnicity)*as.factor(gender) +
                          as.factor(abscanner_1) +
                          agenew + 
                          as.factor(married) +
                          edu + 
                          income + 
                          as.factor(hasnoinsurance) +
                          sbpavg +
                          dbpavg + 
                          choltot + 
                          a1c + 
                          as.factor(smkcur) +
                          as.factor(apoe4_positivity),
                     data = withabv1_nl)
summary(model2_amyloid_nl)

## outputting results
model2_amyloid_nl_est <- cbind(coef = coef(model2_amyloid_nl),
                            confint(model2_amyloid_nl, level = 0.95),
                            pvalue = summary(model2_amyloid_nl)$coefficients[,4]) %>% 
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "amyloid")
writeDataTable(cognl_wb,
               sheet = "amyloid",
               x = model2_amyloid_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)
```
## Model 2 stratified by gender
```{r amyloid model 2 by gender}
## men
model2men_amyloid_nl <- lm(scale(absuvr_1) ~
                          as.factor(ethnicity) +
                          as.factor(abscanner_1) +
                          agenew +
                          edu + 
                          income + 
                          as.factor(hasnoinsurance) + 
                          as.factor(married) +
                          sbpavg +
                          dbpavg + 
                          choltot + 
                          a1c + 
                          as.factor(smkcur) +
                          as.factor(apoe4_positivity),
                        data = menwithabv_nl)
summary(model2men_amyloid_nl)

## outputting results
model2men_amyloid_nl_est <- cbind(coef = coef(model2men_amyloid_nl),
                               confint(model2men_amyloid_nl, level = 0.95), 
                               pvalue = summary(model2men_amyloid_nl)$coefficients[,4]) %>% 
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "amyloid men")
writeDataTable(cognl_wb,
               sheet = "amyloid men",
               x = model2men_amyloid_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## women
model2women_amyloid_nl <- lm(scale(absuvr_1) ~
                            as.factor(ethnicity) +
                            as.factor(abscanner_1) +
                            agenew +
                            edu +
                            income +
                            as.factor(hasnoinsurance) +
                            as.factor(married) +
                            sbpavg +
                            dbpavg + 
                            choltot + 
                            a1c + 
                            as.factor(smkcur) +
                          as.factor(apoe4_positivity),
                          data = womenwithabv_nl)
summary(model2women_amyloid_nl)

## outputting results
model2women_amyloid_nl_est <- cbind(coef = coef(model2women_amyloid_nl),
                                 confint(model2women_amyloid_nl, level = 0.95),
                                 pvalue = summary(model2women_amyloid_nl)$coefficients[,4]) %>%
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "amyloid women")
writeDataTable(cognl_wb,
               sheet = "amyloid women",
               x = model2women_amyloid_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)
```
```{r amyloid clean, include = FALSE}
## cleaning up workspace
rm(model2_amyloid_nl, model2_amyloid_nl_est,
   model2men_amyloid_nl, model2men_amyloid_nl_est,
   model2women_amyloid_nl, model2women_amyloid_nl_est,
   menwithabv_nl, womenwithabv_nl)
```
# Models for tau medial temporal deposition 
Note that participants were scanned with the same scanner, so adding that variable is not needed. 

## Creating gender stratified datasets
```{r mttau strat, include = FALSE}
menwithmttau_nl <- withmttau1_nl %>% filter(gender == 0) # n=222
womenwithmttau_nl <- withmttau1_nl %>% filter(gender == 1) # n=428
```
## Model 2 with interaction term
```{r mttau model 2 with intx, include = FALSE}
# Model 2: Model 1 + SBP + DBP + total chol + a1c + current smoker 
model2_mttau_nl <- lm(scale(taumedtempsuvr_1) ~
                        as.factor(ethnicity) +
                        as.factor(gender) +
                        as.factor(ethnicity)*as.factor(gender) +
                        agenew + 
                        as.factor(married) +
                        edu + 
                        income + 
                        as.factor(hasnoinsurance) +
                        sbpavg +
                        dbpavg + 
                        choltot + 
                        a1c + 
                        as.factor(smkcur) +
                        as.factor(apoe4_positivity),
                   #as.factor(tauscanner_1), did not include as participants here were scanned with the same scanner
                   data = withmttau1_nl)
summary(model2_mttau_nl)

## outputting results
model2_mttau_nl_est <- cbind(coef = coef(model2_mttau_nl),
                          confint(model2_mttau_nl, level = 0.95), 
                          pvalue = summary(model2_mttau_nl)$coefficients[,4]) %>%
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "mttau")
writeDataTable(cognl_wb,
               sheet = "mttau",
               x = model2_mttau_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)

```
## Model 2 stratified by gender
```{r mttau model 2 by gender}
# Model 2: Model 1 + SBP + DBP + total chol + a1c + current smoker 
## men
model2men_mttau_nl <- lm(scale(taumedtempsuvr_1) ~
                           as.factor(ethnicity) +
                           agenew +
                           edu + 
                           income + 
                           as.factor(hasnoinsurance) + 
                           as.factor(married) +
                           sbpavg +
                           dbpavg + 
                           choltot + 
                           a1c + 
                           as.factor(smkcur) +
                           as.factor(apoe4_positivity),
                      data = menwithmttau_nl)
summary(model2men_mttmodel2men_mttau_nlau)

## outputting results
model2men_mttau_nl_est <- cbind(coef = coef(model2men_mttau_nl),
                             confint(model2men_mttau_nl, level = 0.95),
                             pvalue = summary(model2men_mttau_nl)$coefficients[,4]) %>%
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "mttau men")
writeDataTable(cognl_wb,
               sheet = "mttau men",
               x = model2men_mttau_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)

## women
model2women_mttau_nl <- lm(scale(taumedtempsuvr_1) ~
                          as.factor(ethnicity) +
                          agenew +
                          edu +
                          income +
                          as.factor(hasnoinsurance) +
                          as.factor(married) +
                          sbpavg +
                          dbpavg + 
                          choltot + 
                          a1c + 
                          as.factor(smkcur) +
                          as.factor(apoe4_positivity),
                        data = womenwithmttau_nl)
summary(model2women_mttau_nl)

## outputting results
model2women_mttau_nl_est <- cbind(coef = coef(model2women_mttau_nl),
                               confint(model2women_mttau_nl, level = 0.95),
                               pvalue = summary(model2women_mttau_nl)$coefficients[,4]) %>%
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "mttau women")
writeDataTable(cognl_wb,
               sheet = "mttau women",
               x = model2women_mttau_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)
```
```{r mttau clean, include = FALSE}
## cleaning up workspace
rm(model2_mttau_nl, model2_mttau_nl_est,
   model2men_mttau_nl, model2men_mttau_nl_est,
   model2women_mttau_nl, model2women_mttau_nl_est,
   menwithmttau_nl, womenwithmttau_nl)
```
# Models for cortical thickness  
## Creating gender stratified datasets
```{r ct strat, include = FALSE}
menwithct_nl <- withctv1_nl %>% filter(gender == 0) # n=620
womenwithct_nl <- withctv1_nl %>% filter(gender == 1) # n=1412
```
## Model 2 with interaction term
```{r ct model 2 with intx, include = FALSE}
model2_ct_nl <- lm(scale(ctmetaroi_1) ~
                     as.factor(ethnicity) +
                     as.factor(gender) +
                     as.factor(ethnicity)*as.factor(gender) +
                     agenew +
                     as.factor(married) +
                     edu + 
                     income + 
                     as.factor(hasnoinsurance) +
                     as.factor(mriscanner_1) +
                     sbpavg +
                     dbpavg + 
                     choltot + 
                     a1c +
                     as.factor(smkcur) +
                     as.factor(apoe4_positivity),
                data = withmttau1_nl)
summary(model2_ct_nl)

## outputting results
model2_ct_nl_est <- cbind(coef = coef(model2_ct_nl),
                       confint(model2_ct_nl, level = 0.95),
                       pvalue = summary(model2_ct_nl)$coefficients[,4]) %>% 
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "ct")
writeDataTable(cognl_wb,
               sheet = "ct",
               x = model2_ct_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)

```
## Model 2 stratified by gender 
```{r ct model 2 by gender, include = FALSE}
# Model 2: Model 1 + SBP + DBP + total chol + a1c + current smoker 
## men
model2men_ct_nl <- lm(scale(ctmetaroi_1) ~
                        as.factor(ethnicity) +
                        agenew +
                        as.factor(married) +
                        edu +
                        income + 
                        as.factor(hasnoinsurance) +
                        as.factor(mriscanner_1) +
                        sbpavg +
                        dbpavg + 
                        choltot + 
                        a1c + 
                        as.factor(smkcur) +
                        as.factor(apoe4_positivity),
                data = menwithct_nl)
summary(model2men_ct_nl)

## outputting results
model2men_ct_nl_est <- cbind(coef = coef(model2men_ct_nl),
                          confint(model2men_ct_nl, level = 0.95),
                          pvalue = summary(model2men_ct_nl)$coefficients[,4]) %>%
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "ct men")
writeDataTable(cognl_wb,
               sheet = "ct men",
               x = model2men_ct_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)

## women
model2women_ct_nl <- lm(scale(ctmetaroi_1) ~
                          as.factor(ethnicity) +
                          agenew +
                          as.factor(married) +
                          edu +
                          income + 
                          as.factor(hasnoinsurance) +
                          as.factor(mriscanner_1) + 
                          sbpavg +
                          dbpavg + 
                          choltot + 
                          a1c + 
                          as.factor(smkcur) +
                          as.factor(apoe4_positivity),
                     data = womenwithct_nl)
summary(model2women_ct_nl)

## outputting results
model2women_ct_nl <- cbind(coef = coef(model2women_ct_nl),
                            confint(model2women_ct_nl, level = 0.95),
                            pvalue = summary(model2women_ct_nl)$coefficients[,4]) %>%
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "ct women")
writeDataTable(cognl_wb,
               sheet = "ct women",
               x = model2women_ct_nl,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)

```
```{r ct clean, include = FALSE}
## cleaning up workspace
rm(model2_ct_nl, model2_ct_nl_est,
   model2men_ct_nl, model2men_ct_nl_est,
   model2women_ct_nl, model2women_ct_nl_est,
   menwithct_nl, womenwithct_nl)
```
# Models for hippocampal volume
## Creating gender stratified datasets
```{r hipp strat, include = FALSE}
menwithhipp_nl <- withhipp1_nl %>% filter(gender == 0) # n=817
womenwithhipp_nl <- withhipp1_nl %>% filter(gender == 1) # n=1609
```
## Model 2 with interaction term
```{r hipp model 2 with intx, include = FALSE}
# Model 2: Model 1 + SBP + DBP + total chol + a1c + current smoker 
model2_hipp_nl <- lm(scale(hippvolnew_1) ~
                       as.factor(ethnicity) +
                       as.factor(gender) +
                       as.factor(ethnicity)*as.factor(gender) +
                       agenew + 
                       as.factor(married) +
                       edu + 
                       income + 
                       as.factor(hasnoinsurance) +
                       as.factor(mriscanner_1) +
                       sbpavg +
                       dbpavg + 
                       choltot + 
                       a1c + 
                       as.factor(smkcur) +
                       icv_1 +
                       as.factor(apoe4_positivity),
                   data = withhipp1_nl)
summary(model2_hipp_nl)

## outputting results
model2_hipp_nl_est <- cbind(coef = coef(model2_hipp_nl),
                         confint(model2_hipp_nl, level = 0.95),
                         pvalue = summary(model2_hipp_nl)$coefficients[,4]) %>% 
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "hipp")
writeDataTable(cognl_wb,
               sheet = "hipp",
               x = model2_hipp_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)

```
## Model 2 stratified by gender
```{r hipp model 2 by gender}
# Model 2: Model 1 + SBP + DBP + total chol + a1c + current smoker 
## men
model2men_hipp_nl <- lm(scale(hippvolnew_1) ~
                          as.factor(ethnicity) +
                          agenew +
                          edu + 
                          income + 
                          as.factor(hasnoinsurance) + 
                          as.factor(married) +
                          as.factor(mriscanner_1) +
                          sbpavg +
                          dbpavg + 
                          choltot + 
                          a1c + 
                          as.factor(smkcur) +
                          icv_1 +
                          as.factor(apoe4_positivity),
                     data = menwithhipp_nl)
summary(model2men_hipp_nl)

## outputting results
model2men_hipp_nl_est <- cbind(coef = coef(model2men_hipp_nl),
                            confint(model2men_hipp_nl, level = 0.95),
                            pvalue = summary(model2men_hipp_nl)$coefficients[,4]) %>%
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "hipp men")
writeDataTable(cognl_wb,
               sheet = "hipp men",
               x = model2men_hipp_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)

## women
model2women_hipp_nl <- lm(scale(hippvolnew_1) ~
                         as.factor(ethnicity) +
                         agenew +
                         edu +
                         income +
                         as.factor(hasnoinsurance) +
                         as.factor(married) +
                         as.factor(mriscanner_1) +
                         sbpavg +
                         dbpavg + 
                         choltot + 
                         a1c + 
                         as.factor(smkcur) +
                         icv_1 +
                          as.factor(apoe4_positivity),
                       data = womenwithhipp_nl)
summary(model2women_hipp_nl)

## outputting results
model2women_hipp_nl_est <- cbind(coef = coef(model2women_hipp_nl),
                              confint(model2women_hipp_nl, level = 0.95),
                              pvalue = summary(model2women_hipp_nl)$coefficients[,4]) %>%
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "hipp women")
writeDataTable(cognl_wb,
               sheet = "hipp women",
               x = model2women_hipp_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)
```
```{r hipp clean, include = FALSE}
## cleaning up workspace
rm(model2_hipp_nl, model2_hipp_nl_est,
   model2women_hipp_nl, model2women_hipp_nl_est,
   model2men_hipp_nl, model2men_hipp_nl_est,
   menwithhipp_nl,  womenwithhipp_nl)
```

# Models for log(wmhv) deposition 
## Creating gender stratified datasets
```{r wmhv strat, include = FALSE}
menwithwmhv_nl <- withwmhv1_nl %>% filter(gender == 0) # n=876
womenwithwmhv_nl <- withwmhv1_nl %>% filter(gender == 1) # n=1706
```
## Model 2 with interaction term
```{r wmhv model 2 with intx, include = FALSE}
# Model 2: Model 1 + SBP + DBP + total chol + a1c + current smoker 
model2_wmhv_nl <- lm(logwmhv1 ~
                       as.factor(ethnicity) +
                       as.factor(gender) +
                       as.factor(ethnicity)*as.factor(gender) +
                       agenew + 
                       as.factor(married) +
                       edu + 
                       income + 
                       as.factor(hasnoinsurance) +
                       as.factor(mriscanner_1) +
                       sbpavg +
                       dbpavg + 
                       choltot + 
                       a1c + 
                       as.factor(smkcur) +
                       icv_1 +
                       as.factor(apoe4_positivity),
                  data = withwmhv1_nl)
summary(model2_wmhv_nl)

## outputting results
model2_wmhv_nl_est <- cbind(coef = exp(coef(model2_wmhv_nl)),
                         exp(confint(model2_wmhv_nl, level = 0.95)),
                         pvalue = summary(model2_wmhv_nl)$coefficients[,4]) %>%
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "wmhv")
writeDataTable(cognl_wb,
               sheet = "wmhv",
               x = model2_wmhv_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)

```
## Model 2 stratified by gender
```{r wmhv model 2 by gender}
## men
model2men_wmhv_nl <- lm(logwmhv1 ~
                          as.factor(ethnicity) +
                          agenew +
                          edu + 
                          income +
                          as.factor(hasnoinsurance) +
                          as.factor(married) +
                          as.factor(mriscanner_1) +
                          sbpavg +
                          dbpavg + 
                          choltot + 
                          a1c + 
                          as.factor(smkcur) +
                          icv_1 +
                          as.factor(apoe4_positivity),
                     data = menwithwmhv_nl)
summary(model2men_wmhv_nl)

## outputting results
model2men_wmhv_nl_est <- cbind(coef = exp(coef(model2men_wmhv_nl)),
                            exp(confint(model2men_wmhv_nl, level = 0.95)),
                              pvalue = summary(model2men_wmhv_nl)$coefficients[,4]) %>% 
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "wmhv men")
writeDataTable(cognl_wb,
               sheet = "wmhv men",
               x = model2men_wmhv_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)

## women
model2women_wmhv_nl <- lm(logwmhv1 ~
                         as.factor(ethnicity) +
                         agenew +
                         edu +
                         income +
                         as.factor(hasnoinsurance) +
                         as.factor(married) +
                         as.factor(mriscanner_1) +
                         sbpavg +
                         dbpavg + 
                         choltot + 
                         a1c + 
                         as.factor(smkcur) +
                         icv_1 +
                          as.factor(apoe4_positivity),
                       data = womenwithwmhv_nl)
summary(model2women_wmhv_nl)

## outputting results
model2women_wmhv_nl_est <- cbind(coef = exp(coef(model2women_wmhv_nl)),
                              exp(confint(model2women_wmhv_nl, level = 0.95)),
                              pvalue = summary(model2women_wmhv_nl)$coefficients[,4]) %>% 
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "wmhv women")
writeDataTable(cognl_wb,
               sheet = "wmhv women",
               x = model2women_wmhv_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)
```
```{r wmhv clean, include = FALSE}
## cleaning up workspace
rm(model2_wmhv_nl, model2_wmhv_nl_est,
   model2men_wmhv_nl, model2men_wmhv_nl_est,
   model2women_wmhv_nl, model2women_wmhv_nl_est,
   menwithwmhv_nl, womenwithwmhv_nl)
```

# Models for lacunar infarcts
This model did not converge well due to small cell sizes. Code here for reference, but did not present in the paper. 
## Creating gender stratified datasets
```{r lac strat, include = FALSE}
menwithlac_nl <- withlac1_nl %>% filter(gender == 0) # n=893
womenwithlac_nl <- withlac1_nl %>% filter(gender == 1) # n=1741
```
## Model 2 with interaction term
```{r lac model 2 with intx, include = FALSE}
# Model 2: Model 1 + SBP + DBP + total chol + a1c + current smoker 
model2_lac_nl <- glm(lacune_1 ~
                       as.factor(ethnicity) +
                       as.factor(gender) +
                       as.factor(ethnicity)*as.factor(gender) +
                       agenew + 
                       as.factor(married) +
                       edu + 
                       income + 
                       as.factor(hasnoinsurance) + 
                       sbpavg +
                       dbpavg + 
                       choltot + 
                       a1c + 
                       as.factor(smkcur) +
                       as.factor(apoe4_positivity),
                    #as.factor(mriscanner_1),
                  family = binomial(link='logit'),
                  data = withlac1_nl)
summary(model2_lac_nl) # this model does not converge due to small sample size

## outputting results
model2_lac_nl_est <- cbind(coef = exp(coef(model2_lac_nl)),
                        exp(confint(model2_lac_nl, level = 0.95)),
                        pvalue = summary(model2_lac_nl)$coefficients[,4]) %>% 
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "lac")
writeDataTable(cognl_wb,
               sheet = "lac",
               x = model2_lac_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)

```
## Model 2 stratified by gender
There are 0 Hispanic men with lacunar infarcts, so I did not re-estimate stratified analyses. We can present the full model with interaction if needed.

```{r lac clean, include = FALSE}
## cleaning up workspace
rm(model2_lac_nl, model2_lac_nl_est,
   menwithlac_nl, menwithlac_nl)
```
# Models for microbleeds
Note that in the lacunar infarcts analysis, all those with lacunar infarcts were scanned on the Vidal scanner, so the fitted probabilities are numerially 0 or 1 depending on the cell. Did not include scanner for these models for that reason. 

## Creating gender stratified datasets
```{r mb strat, include = FALSE}
menwithmb_nl <- withmb1_nl %>% filter(gender == 0) # n=1239
womenwithmb_nl <- withmb1_nl %>% filter(gender == 1) # n=2140
```

## Model 2 with interaction term
```{r mb model 2 with intx, include = FALSE}
model2_mb_nl <- glm(microhem_1 ~
                      as.factor(ethnicity) +
                      as.factor(gender) +
                      as.factor(ethnicity)*as.factor(gender) +
                      agenew + 
                      as.factor(married) +
                      edu + 
                      income + 
                      as.factor(hasnoinsurance) +
                      sbpavg +
                      dbpavg + 
                      choltot + 
                      a1c + 
                      as.factor(smkcur) +
                      as.factor(apoe4_positivity),
                 #as.factor(mriscanner_1),
                 family = binomial(link='logit'),
                 data = withmb1_nl)
summary(model2_mb_nl)

## outputting results
model2_mb_nl_est <- cbind(coef = exp(coef(model2_mb_nl)),
                        exp(confint(model2_mb_nl, level = 0.95)),
                       pvalue = summary(model2_mb_nl)$coefficients[,4]) %>%
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "mb")
writeDataTable(cognl_wb,
               sheet = "mb",
               x = model2_mb_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)

```
## Model 2 stratified by gender
```{r mb model 2 by gender}
## men
model2men_mb_nl <- glm(microhem_1 ~
                         as.factor(ethnicity) +
                         agenew +
                         edu + 
                         income + 
                         as.factor(hasnoinsurance) + 
                         as.factor(married) +
                         #as.factor(mriscanner_1) +
                         sbpavg +
                         dbpavg + 
                         choltot + 
                         a1c + 
                         as.factor(smkcur) +
                         as.factor(apoe4_positivity),
                     family = binomial(link='logit'),
                     data = menwithmb_nl)
summary(model2men_mb_nl)

## outputting results
model2men_mb_nl_est <- cbind(coef = exp(coef(model2men_mb_nl)),
                          exp(confint(model2men_mb_nl, level = 0.95)),
                          pvalue = summary(model2men_mb_nl)$coefficients[,4]) %>%
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "mb men")
writeDataTable(cognl_wb,
               sheet = "mb men",
               x = model2men_mb_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)

## women
model2women_mb_nl <- glm(microhem_1 ~
                         as.factor(ethnicity) +
                         agenew +
                         edu +
                         income +
                         as.factor(hasnoinsurance) +
                         as.factor(married) +
                         #as.factor(mriscanner_1) +
                         sbpavg +
                         dbpavg + 
                         choltot + 
                         a1c + 
                         as.factor(smkcur) +
                          as.factor(apoe4_positivity),
                       family = binomial(link='logit'),
                       data = womenwithmb_nl)
summary(model2women_mb_nl)

## outputting results
model2women_mb_nl_est <- cbind(coef = exp(coef(model2women_mb_nl)),
                            exp(confint(model2women_mb_nl, level = 0.95)),
                            pvalue = summary(model2women_mb_nl)$coefficients[,4]) %>%
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(cognl_wb, sheetName = "mb women")
writeDataTable(cognl_wb,
               sheet = "mb women",
               x = model2women_mb_nl_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(cognl_wb,
             "./Results - RaceEthGenImg/cognl.xlsx",
             overwrite = TRUE)
```

```{r mb clean, include = FALSE}
## cleaning up workspace
rm(model2_mb_nl, model2_mb_nl_est,
   model2men_mb_nl, model2men_mb_nl_est,
   model2women_mb_nl, model2women_mb_nl_est)
```