---
title: "4b. supp_RaceEthGenImg_IPSW_02 14 25"
output: html_document
date: "2025-02-14"
---
# Purpose of code
To re-weight the models for PET outcomes for selection into the PET subsample. 

# Loading packages
```{r packagenews, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(haven) # inputting data 
library(tidyverse) # data mgmt
library(psych) # easy summary statistics
library(DT) # data tables
library(tableone) # easy table 1 
library(kableExtra) # format kable objects 
library(naniar) # for missingness 
library(openxlsx) # excel for outputting results 
library(survey) # survey glm
library(broom) # making output more readable 
options(max.print=100000)

# turn off scientific notation
options(scipen = 999)
```
# Loading data
```{r data, include = FALSE}
load("./aim2_final.RData") # n=3840, v=338
view(names(aim2_final_frs))

# those dementia-free at baseline with any imaging data at v1
withimg <- aim2_final %>%
  filter(!is.na(cdx_cog) & !cdx_cog == 2) %>% # excluding those w/ dementia
  filter( (hasabv1 == 1) | (hasctv1 == 1) | (haswmhv1 == 1) | (hasmttauv1 == 1) | (hashipp1 == 1)) 
 # n=3433, v=338 

# how many people are in amyloid subsample?
summary(as.factor(withimg$hasabv1)) # 1849 excluded

# how many people are in tau subsample?
summary(as.factor(withimg$hasmttauv1)) # 2519 excluded
```
# Creating excel sheet for results
```{r amyloid excel, include = FALSE}
# if starting over and re-running all models, run this
ipw_wb <- createWorkbook()

# if coming back to analysis, run this
ipw_wb <- loadWorkbook("./Results - RaceEthGenImg/ipw.xlsx")
```
# Creating weights
Creating stabilized IPW weights for selection into the amyloid and tau subsamples. Will not include covariates here and just include in the main models so that we can keep the weights as simple as possible.
```{r create ipw, include = FALSE}
# amyloid subsample
## denominator model
den_fit_ab <- glm(hasabv1 ~ 
                    agenew +
                    as.factor(ethnicity) +
                    as.factor(gender),
                  family = binomial(link = "logit"),
                  data = withimg)
summary(den_fit_ab)

## numerator model
num_fit_ab <- glm(hasabv1 ~ 1,
                  family = binomial(link = "logit"),
                  data = withimg)
summary(num_fit_ab)

## denominator propensities
den_pred_ab <- predict(den_fit_ab, type = "response")
summary(den_pred_ab)

## numerator propensities
num_pred_ab <- predict(num_fit_ab, type = "response")
summary(num_pred_ab)

## creating weights 
sw_ab <- ifelse(withimg$hasabv1 == 1, 
                num_pred_ab/den_pred_ab, # weights for those in the sample
                (1-num_pred_ab)/(1-den_pred_ab)) # weights for those not in the sample

# tau subsample
## denominator model
den_fit_tau <- glm(hasmttauv1 ~ 
                     agenew +
                     as.factor(ethnicity) +
                     as.factor(gender),
                   family = binomial(link = "logit"),
                   data = withimg)
summary(den_fit_tau)

## numerator model
num_fit_tau <- glm(hasmttauv1 ~ 1,
                   family = binomial(link = "logit"),
                   data = withimg)
summary(num_fit_tau)

## denominator propensities
den_pred_tau <- predict(den_fit_tau, type = "response")
summary(den_pred_tau)

## numerator propensities
num_pred_tau <- predict(num_fit_tau, type = "response")
summary(num_pred_tau)

## creating weights 
sw_tau <- ifelse(withimg$hasmttauv1 == 1, 
                 num_pred_tau/den_pred_tau, # weights for those in the sample
                 (1-num_pred_tau)/(1-den_pred_tau)) # weights for those not in the sample
summary(sw_tau) 

# adding weights to dataset
withimg_sw <- bind_cols(withimg, 
                        sw_amyloid = sw_ab, 
                        sw_mttau = sw_tau) # n=3433, v=62
view(names(withimg_sw))

## checking weights
describeBy(withimg_sw$sw_amyloid, withimg_sw$hasabv1)
describeBy(withimg_sw$sw_mttau, withimg_sw$hasmttauv1)
```
# Creating gender stratified datasets
```{r gen strat data, include = FALSE}
# men
menwithimg_sw <- withimg_sw %>%
  filter(gender == 0) #n=1261, v=362

# women
womenwithimg_sw <- withimg_sw %>%
  filter(gender == 1) # n=2172, v=362
```
# Running analyses with weights 
## Amyloid
```{r amyloid ipw, include = FALSE}
design_ab <- svydesign(ids=~1, weights=~sw_amyloid, data=withimg_sw)

# Model 2 with interaction terms
amyloid_ipw <- svyglm(scale(absuvr_1) ~
                        as.factor(ethnicity) +
                        as.factor(gender) +
                        as.factor(ethnicity)*as.factor(gender) +
                        as.factor(abscanner_1) +
                        agenew + 
                        as.factor(married) +
                        edu + 
                        income + 
                        as.factor(hasnoinsurance) +
                        sbpavg +
                        dbpavg + 
                        choltot + 
                        a1c + 
                        as.factor(smkcur) +
                        as.factor(apoe4_positivity),
                   data = withimg_sw,
                   design = design_ab)
summary(amyloid_ipw)

## outputting results
amyloid_ipw_est <- cbind(coef = coef(amyloid_ipw),
                         confint(amyloid_ipw, level = 0.95),
                         pvalue = summary(amyloid_ipw)$coefficients[,4]) %>% 
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(ipw_wb, sheetName = "amyloid ipw")
writeDataTable(ipw_wb,
               sheet = "amyloid ipw",
               x = amyloid_ipw_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(ipw_wb,
             "./Results - RaceEthGenImg/ipw.xlsx",
             overwrite = TRUE)

## Model 2 stratified by gender 
## men
design_men_ab <- svydesign(ids=~1, weights=~sw_amyloid, data=menwithimg_sw)
amyloid_men_ipw <- svyglm(scale(absuvr_1) ~
                            as.factor(ethnicity) +
                            as.factor(abscanner_1) +
                            agenew + 
                            as.factor(married) +
                            edu + 
                            income + 
                            as.factor(hasnoinsurance) +
                            sbpavg +
                            dbpavg + 
                            choltot + 
                            a1c + 
                            as.factor(smkcur) +
                            as.factor(apoe4_positivity),
                          data = menwithimg_sw,
                          design = design_men_ab)
summary(amyloid_men_ipw)

## outputting results
amyloid_men_ipw_est <- cbind(coef = coef(amyloid_men_ipw),
                             confint(amyloid_men_ipw, level = 0.95), 
                             pvalue = summary(amyloid_men_ipw)$coefficients[,4]) %>% 
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(ipw_wb, sheetName = "amyloid ipw men")
writeDataTable(ipw_wb,
               sheet = "amyloid ipw men",
               x = amyloid_men_ipw_est,
               colNames = TRUE,
               rowNames = TRUE)

## women
design_women_ab <- svydesign(ids=~1, weights=~sw_amyloid, data=womenwithimg_sw)
amyloid_women_ipw <- svyglm(scale(absuvr_1) ~
                              as.factor(ethnicity) +
                              as.factor(abscanner_1) +
                              agenew + 
                              as.factor(married) +
                              edu + 
                              income + 
                              as.factor(hasnoinsurance) +
                              sbpavg +
                              dbpavg + 
                              choltot + 
                              a1c + 
                              as.factor(smkcur) +
                              as.factor(apoe4_positivity),
                          data = womenwithimg_sw,
                          design = design_women_ab)
summary(amyloid_women_ipw)

## outputting results
amyloid_women_ipw_est <- cbind(coef = coef(amyloid_women_ipw),
                               confint(amyloid_women_ipw, level = 0.95),
                               pvalue = summary(amyloid_women_ipw)$coefficients[,4]) %>%
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(ipw_wb, sheetName = "amyloid ipw women")
writeDataTable(ipw_wb,
               sheet = "amyloid ipw women",
               x = amyloid_women_ipw_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(ipw_wb,
             "./Results - RaceEthGenImg/ipw.xlsx",
             overwrite = TRUE)
```
## Tau
```{r tau ipw, include = FALSE}
# Model 2 with interaction terms
design_tau <- svydesign(ids=~1, weights=~sw_mttau, data=withimg_sw)

mttau_ipw <- svyglm(scale(taumedtempsuvr_1) ~
                      as.factor(ethnicity) +
                      as.factor(gender) +
                      as.factor(ethnicity)*as.factor(gender) +
                      agenew + 
                      as.factor(married) +
                      edu + 
                      income + 
                      as.factor(hasnoinsurance) +
                      sbpavg +
                      dbpavg + 
                      choltot + 
                      a1c + 
                      as.factor(smkcur) +
                      as.factor(apoe4_positivity),
                    data = withimg_sw,
                    design = design_tau)
summary(mttau_ipw)

## outputting results
mttau_ipw_est <- cbind(coef = coef(mttau_ipw),
                       confint(mttau_ipw, level = 0.95),
                       pvalue = summary(mttau_ipw)$coefficients[,4]) %>% 
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(ipw_wb, sheetName = "mttau ipw")
writeDataTable(ipw_wb,
               sheet = "mttau ipw",
               x = mttau_ipw_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(ipw_wb,
             "./Results - RaceEthGenImg/ipw.xlsx",
             overwrite = TRUE)

## Model 2 stratified by gender 
## men
design_men_tau <- svydesign(ids=~1, weights=~sw_mttau, data=menwithimg_sw)

mttau_men_ipw <- svyglm(scale(taumedtempsuvr_1) ~
                          as.factor(ethnicity) +
                          agenew +
                          as.factor(married) +
                          edu + 
                          income + 
                          as.factor(hasnoinsurance) +
                          sbpavg +
                          dbpavg +
                          choltot +
                          a1c + 
                          as.factor(smkcur) +
                          as.factor(apoe4_positivity),
                        data = menwithimg_sw,
                        design = design_men_tau)
summary(mttau_men_ipw)

## outputting results
mttau_men_ipw_est <- cbind(coef = coef(mttau_men_ipw),
                           confint(mttau_men_ipw, level = 0.95), 
                           pvalue = summary(mttau_men_ipw)$coefficients[,4]) %>% 
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(ipw_wb, sheetName = "mttau ipw men")
writeDataTable(ipw_wb,
               sheet = "mttau ipw men",
               x = mttau_men_ipw_est,
               colNames = TRUE,
               rowNames = TRUE)

## women
design_women_tau <- svydesign(ids=~1, weights=~sw_mttau, data=womenwithimg_sw)
mttau_women_ipw <- svyglm(scale(taumedtempsuvr_1) ~
                            as.factor(ethnicity) +
                            agenew + 
                            as.factor(married) +
                            edu + 
                            income + 
                            as.factor(hasnoinsurance) +
                            sbpavg +
                            dbpavg + 
                            choltot + 
                            a1c + 
                            as.factor(smkcur) +
                            as.factor(apoe4_positivity),
                          data = womenwithimg_sw,
                          design = design_women_tau)
summary(mttau_women_ipw)


## outputting results
mttau_women_ipw_est <- cbind(coef = coef(mttau_women_ipw),
                             confint(mttau_women_ipw, level = 0.95),
                             pvalue = summary(mttau_women_ipw)$coefficients[,4]) %>%
  round(., digits = 2) %>%
  as.data.frame() %>% view()

addWorksheet(ipw_wb, sheetName = "mttau ipw women")
writeDataTable(ipw_wb,
               sheet = "mttau ipw women",
               x = mttau_women_ipw_est,
               colNames = TRUE,
               rowNames = TRUE)

## saving workbook
saveWorkbook(ipw_wb,
             "./Results - RaceEthGenImg/ipw.xlsx",
             overwrite = TRUE)
```

